SoK: Defending Against Web Application Attacks: From Research Approaches to Practical Tools

### 2014-11-15 docs/questioning.tex
pass in VI

In this paper, we explore how different attacks associated with 
the exploitation of untrusted input validation errors can be modeled 
under a {\it common perspective}. To that end, we propose an exploitation
model which highlights that most of the steps needed to mount
different types of code injection attacks are
common.\footnote{Note that we do not consider lower-level attacks based on
the exploitation of memory corruption vulnerabilities and the injection
of binary code---see Section~\ref{sec:attacks}.} This is validated by the fact that
there are protection mechanisms that can defend against more than one of these
types of attacks. Then, we {\it categorize} a selection of representative
protection mechanisms and finally we {\it analyze} each mechanism across the
following dimensions:
\begin{itemize}
\item {\it Accuracy:} protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability:} whether the mechanism and its
  testbed are publicly available.
\item {\it Computational performance:} the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use:} whether the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security:} the robustness of the protection mechanism against
  attackers with knowledge of its internals who attempt to circumvent it.
\item {\it Detection Point:} the location where a mechanism detects an attack
  based on our exploitation model.
\end{itemize}

@@

The first two criteria correspond to \emph{Accuracy} and
\emph{Availability}. The other four criteria correspond to \emph{Computational
  Performance}, \emph{Ease of Use}, \emph{Security}, and
\emph{Point of Detection}.

@@

\subsection{Computational Performance}

Detection mechanisms may impose a cost due to their use, as they
typically introduce some amount of extra computation on a existing applications.
The overhead depends on the specifics of each mechanism. For
example, it may be due to some form of run-time checking, or some form of
obfuscation. Also, depending on the approach, the cost may be incurred
on different places: it may affect a server
(e.g., its memory or {\sc cpu} usage, processing throughput, or response latency),
it may affect the client, or both. The
usefulness of a detection mechanism depends therefore on the
computational cost it requires and on where it imposes it, as
different overheads may be acceptable at the server side than the client side.

@@

  & \bf{Computational Overhead}\tnote{3} & \\

@@

@@

    \multicolumn{5}{p{13cm}}{%
  \begin{footnotesize}
       $^\textsuperscript{1}$The check mark (\tick) indicates that the publication
       includes a homepage where the reader can 
       download the corresponding software. {\bf AO} (Available On-line) suggests
       that the software is available on-line but the
       address was not mentioned in the paper, which probably indicates that
       the authors made it available after the publication. The question 
       mark ({\bf ?}) indicates that the a homepage for the software
       was included
       in the publication but now it is not available.
  \end{footnotesize} } \\
    \end{tabular}}
    \end{small}
    \end{threeparttable}
\end{table*}

We group our key observations on the~35 publications that
we considered in this work along the
dimensions we identified in Section~\ref{sec:dimensions}.

Table~\ref{tab:comp} indicates that the authors of 3 out of 35 (8.5\%)
publications provided complete quadruples of {\sc tp}, {\sc tn}, 
{\sc fp}, and {\sc fn}. On the other hand, we
see that 4 out of 35 (11.4\%) were not tested at all.
In these cases, all the four elements of the quadruples
are not available ({\sc na}). An interesting observation is
that in 10 out of the 35 (28.5\%)
publications, the authors performed only attacks during their tests
and did not test their mechanisms to see if they produce any false
alarms. The corresponding quadruples contain {\sc tp} and {\sc fn}
results but {\sc tn} and {\sc fp} results are not available ({\sc
  na}). This stresses the fact that authors may be more interested in
making sure that their mechanisms can detect known attacks rather than
seeing how they respond under normal conditions. In some occasions,
where the number of {\sc tp} results are not quantified, authors did
not mention how many or which attacks they performed. Moreover in some
cases we observe that even if the authors report the existence of
possible false positive and negative results, they have not quantified
them ({\sc nq}).

Even when some numbers are given, they may not be adequate.
Sensitivity and specificity are statistical measures, and as such
should be interpreted with suitable confidence intervals. There are
various methods to calculate confidence intervals, even dispensing
with the need to have large samples in order to be able to use the
central limit theorem~\cite{agresti1998,brown2001}.
However, sample sizes in the single
digits are not enough to produce good intervals.

Regarding the subjects of tests, we see that 27 out of 35 (77.1\%)
mechanisms were tested on real-world applications known to be
vulnerable. The vulnerabilities associated with these applications
are enlisted in the following providers: the Common
Vulnerabilities and Exposures ({\sc cve}) database~\cite{cve},
the Bugtraq security mailing list~\cite{bugtraq}, the {\sc xss}ed.com
security bulletin provider~\cite{xssed}, Microsoft's security
bulletin~\cite{microsoftBulletin} and the
ha.ckers.org security bulletin provider~\cite{hackers}.
Figure~\ref{fig:defect_sources} presents how many, in percentage,
and which, in the figure's caption, authors referred
to which source. In numerous occassions authors performed tests based
on the same applications. For example, 12 mechanisms were tested on a
vulnerable version of {\sc phpbb}.\footnote{https://www.phpbb.com/}

\begin{figure}[t]
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{defect-percentages.png}
\end{center}
\caption{\label{fig:defect_sources}Pie chart showing the sources
that the used to find vulnerabilities to test
their mechanisms. We also provide which source was used
in which publication:
{\sc cve}~\cite{XBS06,NLC07,PMP11,BK04,BV08,JB07,SMS13,WPLKK09,JKK06a,PS11},
{\it Bugtraq}~\cite{PB05,KKVJ06,JEP08},
{\sc xss}ed.com~\cite{NSS06,APKLM10},
ha.ckers.org~\cite{TNH07,PSC09,LV09},
Microsoft~\cite{RDWDE07}.}
\end{figure}

There are also authors that took a different approach.
For instance, during their initial tests,
Stock et al.~\cite{SLMS14} managed to bypass
the browser-based {\sc xss} filters of the
73\% of 1,602 real-world {\sc dom}-based {\sc xss} vulnerabilities.
These vulnerabilities were actually found during
their previous research~\cite{LSJ13}.
Finally, in the publications that
{\sc amnesia} and {\sc sd}river were presented,
authors managed to break existing application suites
and then test the accuracy of their tools on them.

Table~\ref{tab:avail} presents our findings regarding the availability
of each mechanism in terms of source code and corresponding
executables. We also examined the availability of the testbeds
mentioned in the paper. We see that only 6 out of 35 (17.1\%) of the
pubications provided a homepage for their mechanism and 2 from these 6
homepages are currently not available. In 5 cases authors made either
the source code or their executables available after their paper got
published. In one case ({\sc csp}) the source was available before the
publication through the Mozilla community.
Regarding the availability of test materials, we see that
only 3 out of 35 (8.5\%) publications have currently their testbeds
available on-line.

\subsection{Computational Performance}

Table~\ref{tab:comp} shows for every mechanism its corresponding
overhead. In addition, it indicates if the overhead is applied to the
server-side ({\sc s}), or the client-side ({\sc c}). In almost half of
the approaches, 16 out of 35 (45.7\%), the overhead is provided as a
percentage, while in other cases, 7 our of 35 (20\%), the authors
provide the time (in terms of ms) that their mechanisms add to the
normal execution time of the protecting entity. Note that in some
cases, the times of normal executions are not provided,
so it is not possible to move from absolute
time measurements to percentage overheads. In 10 cases (28.5\%), the
overhead was either not available ({\sc na}) or not quantified ({\sc
  nq}). In the case of {\sc php aspis}, its authors indicate that with
their mechanism, the execution time is doubled.

@@

Mechanisms in different categories face different deployment issues.

Consider the majority of mechanisms coming from the policy enforcement
subcategory. In most cases developers should modify multiple
components to use each mechanism. Specifically, mechanisms like
BrowserShield and {\sc beep} require modifications both on the server
and the client side. Thus, it would be difficult for them to be
adopted by both browser vendors and application developers. On the
other hand, there are cases where the policies introduced in the
server are enforced on the client-side, via a library embedded in the
server's response (i.e., Blueprint). Such an approach is convenient
since no modifications are needed on the client.

Extensive modifications in the application's source
code can also be a reason that can make a mechanism
difficult to use. {\sc sql}rand, {\sc amnesia},
mechanisms coming from the parse-tree validation
subcategory and mechanisms coming from the taint
tracking subcategory are such examples.
In the first three cases, programmers should modify every
code fragment that involves the execution of a query.
In the latter case they should also change
all the code fragments that involve user-input handling.

Finally, even though many of the mechanisms that involve
training are easy to deploy, they have a distinct disadvantage.
When the application is altered, mechanisms like {\sc sd}river
and {\sc xss-guard} require a new training phase is necessary.
However, with the increased adoption of automated testing
and continuous integration frameworks, this phase could be
easily repeated.

@@

While describing the various mechanisms in Section~\ref{sec:defs}, we
observed that some can be bypassed by attackers that know how they
work. Still, some mechanisms are designed in such a way that they
can be extended and become immune to the attacks that they are
currently vulnerable to. For example, policy enforcement frameworks
that are based on JavaScript or {\sc html} rewriting can be extended
to detect attacks that leverage {\tt iframe} tags. This does not apply
to all mechanisms though. In particular, training mechanisms like {\sc
  didafit} and {\sc xss-guard} must be redesigned to detect the
attacks we described in the related sections.

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms}---Point {\sc db}.
The other 11 wrap {\sc api} calls related to output
vectors---Point {\sc e}i{\sc q}.

In frameworks that deal with {\sc xss} attacks we
see that attack detection may take place either on the server-side or
on the client-side. In the first case, a proxy is placed on the server
to examine the server's responses before they reach the user's web
browser. In the second case a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 14 out of 24
frameworks (58,3\%) detect {\sc xss} attacks on the
client's side. Note that the majority of the mechanisms that detect
such attacks on the server-side can also detect {\sc dsl} code
injection attacks (6 out of 10).

@@

Our observations lead to some lessons and recommendations that
developers of new mechanisms may find helpful. In particular, they
call for improving accuracy tests and code availability, while
aiming to reduce computational and deployment overheads.

@@

\subsection{Computational Overhead and Deployment Remarks}

Computational overhead comes up as a non-trivial issue. We observe
that the average overhead that can be extracted from the publications
that provide the overhead in percentage terms is 17.45\%. Relative
research on protection mechanisms~\cite{SPWS13} indicates that
mechanisms that introduce an overhead larger than 10\% do not tend to
gain wide adoption in production environments. Hence, the
computational overhead of some mechanisms could be a reason why they
have not been adopted.

>>>>

In this paper, we explore how different attacks associated with 
the exploitation of untrusted input validation errors can be modeled 
under a {\it common perspective}. To that end, we propose an exploitation
model which highlights that most of the steps needed to mount
different types of code injection attacks are
common.\footnote{Note that we do not consider lower-level attacks based on
the exploitation of memory corruption vulnerabilities and the injection
of binary code---see Section~\ref{sec:attacks}.} This is validated by the fact that
there are protection mechanisms that can defend against more than one of these
types of attacks. Then, we {\it categorize} a selection of representative
protection mechanisms and finally we {\it analyze} each mechanism across the
following dimensions:
\begin{itemize}
\item {\it Accuracy:} protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability:} whether the mechanism and its
  testbed are publicly available.
\item {\it Performance overhead:} the 
  overhead imposed by the mechanisms at their points of deployment.
\item {\it Ease of use:} whether the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security:} the robustness of the protection mechanism against
  attackers with knowledge of its internals who attempt to circumvent it.
\item {\it Detection Point:} the location where a mechanism detects an attack
  based on our exploitation model.
\end{itemize}

@@

The first two criteria correspond to \emph{Accuracy} and
\emph{Availability}. The other four criteria correspond to \emph{Runtime
  Performance}, \emph{Ease of Use}, \emph{Security}, and
\emph{Point of Detection}.

@@

\subsection{Performance Overhead}

Detection mechanisms may impose a cost due to their use, as they
typically introduce some amount of extra computation on existing applications.
The overhead depends on the specifics of each mechanism. For
example, it may be due to some form of run-time checking, or some form of
obfuscation. Also, depending on the approach, the cost may be incurred
on different places: it may affect a server
(e.g., its memory or {\sc cpu} usage, processing throughput, or response latency),
it may affect the client, or both. The
usefulness of a detection mechanism depends therefore on the
computational cost it requires and on where it imposes it, as
different overheads may be acceptable at the server side than the client side.

@@

  & \bf{Performance Overhead}\tnote{3} & \\

@@

We group our key observations on the~35 publications that
we consider in this work along the
dimensions we identified in Section~\ref{sec:dimensions}.

Table~\ref{tab:comp} indicates that the authors of 3 out of 35 (8.5\%)
publications provided a complete tuple of {\sc tp}, {\sc tn}, 
{\sc fp}, and {\sc fn}. On the other hand, we
see that 4 out of 35 (11.4\%) were not tested at all.
In these cases, all four elements of the tuple
are not available ({\sc na}). An interesting observation is
that in 10 out of the 35 (28.5\%)
publications, the evaluation focused on only on attack detection,
and there was no test focusing on false positives.
The corresponding tuples contain {\sc tp} and {\sc fn}
results, but {\sc tn} and {\sc fp} results are not available ({\sc
  na}). This stresses the fact that authors may be more interested in
making sure that their mechanisms can detect known attacks rather than
seeing how they respond under normal conditions. In some occasions,
where the number of {\sc tp} results are not quantified, authors did
not mention how many or which attacks they performed. Moreover, in some
cases we observe that even if the authors report the existence of
possible false positive and negative results, they have not quantified
them ({\sc nq}).

Even when some numbers are given, they may not be adequate.
Sensitivity and specificity are statistical measures, and as such,
they should be interpreted with suitable confidence intervals. There are
various methods to calculate confidence intervals, even without
the need to have large sets of samples in order to be able to use the
central limit theorem~\cite{agresti1998,brown2001}.
However, sample sizes in the single
digits are not enough to produce good intervals.

Regarding the subjects of tests, we see that 27 out of 35 (77.1\%)
mechanisms were tested with real-world applications known to be
vulnerable. The vulnerabilities associated with these applications
are enlisted in the following providers: the Common
Vulnerabilities and Exposures ({\sc cve}) database~\cite{cve},
the Bugtraq security mailing list~\cite{bugtraq}, the {\sc xss}ed.com
security bulletin provider~\cite{xssed}, Microsoft's security
bulletin~\cite{microsoftBulletin} and the
ha.ckers.org security bulletin provider~\cite{hackers}.
Figure~\ref{fig:defect_sources} presents how many, in percentage,
and which, in the figure's caption, publications referred
to which source. In numerous occasions authors performed tests based
on the same applications. For example, 12 mechanisms were tested on a
vulnerable version of the {\sc phpbb} bulletin board software.

\begin{figure}[t]
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{defect-percentages.png}
\end{center}
\caption{\label{fig:defect_sources}Most common sources of real-world vulnerabilities
used for testing of web attack defenses.
Specifically, the publications that used a particular source are as follows:
{\sc cve}~\cite{XBS06,NLC07,PMP11,BK04,BV08,JB07,SMS13,WPLKK09,JKK06a,PS11},
{\it Bugtraq}~\cite{PB05,KKVJ06,JEP08},
{\sc xss}ed.com~\cite{NSS06,APKLM10},
ha.ckers.org~\cite{TNH07,PSC09,LV09},
Microsoft~\cite{RDWDE07}.}
\end{figure}

There are also authors that took a different approach.
For instance, during their initial tests,
Stock et al.~\cite{SLMS14} managed to bypass
the browser-based {\sc xss} filters of
73\% out of 1,602 real-world {\sc dom}-based {\sc xss} vulnerabilities.
These vulnerabilities were actually found as part of
their previous research~\cite{LSJ13}.
Finally, in the {\sc amnesia}~\cite{HO05,HO06,HO05b} and
{\sc sd}river~\cite{MS09,MKS09,MKLS11} publications,
the authors managed to break existing application suites
and then test the accuracy of their tools on them.

@@

    \multicolumn{5}{p{12.4cm}}{%
  \begin{footnotesize}
       $^\textsuperscript{1}$A check mark (\tick) indicates that the publication
       includes a link to a page where the software is available.
       {\bf AO} (Available On-line) indicates
       that the software is available on-line but the
       address is not mentioned in the paper, which probably means that
       the it was made available after the publication. A question 
       mark ({\bf ?}) indicates that a link to the software
       was included in the publication but is now inaccessible.
  \end{footnotesize} } \\
    \end{tabular}}
    \end{small}
    \end{threeparttable}
\end{table*}

Table~\ref{tab:avail} presents our findings regarding the availability
of each mechanism in terms of source code and corresponding
executables. We also examined the availability of the testbeds
mentioned in each paper. We see that only 6 out of 35 (17.1\%) of the
publications provided a link to their mechanism and, 2 from these 6
web pages are currently not available. In 5 cases authors made either
the source code or their executables available after their paper got
published. In one case ({\sc csp}), the source was available before the
publication through the Mozilla community.
Regarding the availability of test materials, we see that
only 3 out of 35 (8.5\%) publications have currently their testbeds
available on-line.

\subsection{Performance Overhead}

Table~\ref{tab:comp} shows the performance overhead of each mechanism.
In addition, it indicates if the overhead is incurred at the
server ({\sc s}) or at the client side ({\sc c}). In almost half of
the approaches, 16 out of 35 (45.7\%), the overhead is provided as a
percentage, while in other cases, 7 our of 35 (20\%), the authors
provide the latency (in ms) that their mechanisms add to the
normal execution time of the protected application. Note that in some
cases, the times of normal executions are not provided,
so it is not possible to convert absolute
time measurements to percentage overheads. In 10 cases (28.5\%), the
overhead was either not available ({\sc na}) or not quantified ({\sc
  nq}). For {\sc php aspis}, its authors indicate that with
their mechanism, the execution time is doubled.

@@

Mechanisms in different categories face different deployment issues.
Consider the majority of mechanisms in the policy enforcement
subcategory. In most cases, developers should modify multiple
components to use each mechanism. Specifically, mechanisms like
BrowserShield and {\sc beep} require modifications both on the server
and at the client. Thus, it would be difficult to be
adopted by both browser vendors and application developers. On the
other hand, there are cases where the policies introduced at the
server are enforced at the client side, via a library embedded in the
server's response (i.e., Blueprint). Such an approach is convenient
as no modifications are needed at the client.

Extensive modifications in the application's source
code can also be a reason that can make a mechanism
difficult to use. {\sc sql}rand, {\sc amnesia},
mechanisms of the parse-tree validation
subcategory, and mechanisms of the taint
tracking subcategory are such examples.
In the first three cases, programmers should modify every
code fragment that involves the execution of a query.
In the latter case they should also change
all the code fragments that involve user input handling.

Finally, even though many of the mechanisms that involve
training are easy to deploy, they have a distinct disadvantage.
When the application is altered, mechanisms like {\sc sd}river
and {\sc xss-guard} require a new training phase.
However, with the increased adoption of automated testing
and continuous integration frameworks, this phase could be
easily repeated.

@@

In our discussion of the various mechanisms in Section~\ref{sec:defs}, we
observed that some of them can be bypassed by attackers that know
the internals of their operation.
Still, the design of some mechanisms allows them to
be extended and become immune to the circumvention attacks they are
currently vulnerable to. For example, policy enforcement frameworks
based on JavaScript or {\sc html} rewriting can be extended
to detect attacks that leverage {\tt iframe} tags. This does not apply
to all mechanisms though. In particular, training mechanisms like {\sc
  didafit} and {\sc xss-guard} must be redesigned to detect the
attacks we described in the related sections.

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
at the server side, all mechanisms that detect
such attacks are placed at some point within the server-side infrastructure.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms} (Point {\sc db}).
The other 11 are based on interposing on {\sc api} calls related to output
vectors (Point {\sc e}i{\sc q}).

For frameworks that deal with {\sc xss} attacks, we
see that attack detection and prevention may take place either at the server or
the client side. In the first case, a proxy is typically placed in front of the server
to examine the server's responses before they reach a user's 
browser. In the second case, a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 14 out of 24
frameworks (58,3\%) detect {\sc xss} attacks at the
client side. Note that the majority of the mechanisms that detect
such attacks on the server can also detect {\sc dsl} code
injection attacks (6 out of 10).

@@

Our observations lead to some lessons and recommendations that
developers of new mechanisms may find helpful. In particular, they
call for improving accuracy tests and code availability, while
aiming to reduce performance and deployment overheads.

@@

\subsection{Performance Overhead and Deployment Remarks}

Performance overhead comes up as a non-trivial issue. We observe
that the average overhead that can be extracted from the publications
that provide the overhead in percentage terms is 17.45\%. Relative
research on protection mechanisms~\cite{SPWS13} indicates that
mechanisms that introduce an overhead larger than 10\% do not tend to
gain wide adoption in production environments. Hence, the
computational overhead of some mechanisms could be a reason why they
have not been adopted.
